---
title: "Explore gapminder"
author: "sukey"
date: "September 20, 2018"
output:
    html_document:
        keep_md: true
---

## Smell test the data
Explore the gapminder object:
```{r,warning=FALSE,message=FALSE}
library(gapminder)
library(tidyverse)
library(dplyr)
```

  * #### Is it a data.frame, a matrix, a vector, a list?
```{r, echo=T,results='hide'}
inherits(gapminder,"data.frame") 
### or we can also use "str()"
str(gapminder)
```
**<u> Answer: The function returns "True", so it is a data frame </u>. Using str() gives you that information plus extra goodies (such as the levels of your factors and the first few values of each variable)** 

  * #### What is its class?
```{r, results='hide'}
class(gapminder) 
```
**<u> Answer: The class of gapminder is  "data.frame" </u>. The function class prints the vector names of classes an object inherits from**  

  * #### How many variables/columns?
```{r,results='hide'}
ncol(gapminder) 
```
**<u> Answer: The number of columns is 6 </u>. Also we can use str() function and then count the number of columns**  

  * #### How many rows/observations?
```{r,results='hide'}
nrow(gapminder) 
```
**<u> Answer: The number of rows is 1704 </u>.**

  * #### Can you get these facts about “extent” or “size” in more than one way? Can you imagine different    functions being useful in different contexts?
```{r,results='hide'}
dim(gapminder) # this function can directly return the size
str(gapminder) # str() returns size and other info
nrow(gapminder) # nrow() can combine with ncol() to check the size of data
ncol(gapminder)
```
**<u> Answer:  1704 rows and 6 columns </u>. dim() returns the dimensions of an object. rowNumber = dim(object)[1]; colNumber = dim(object)[2]**  

  * ####  What data type is each variable?
```{r,results='hold'}
str(gapminder)
```
**<u> Answer: The type of each variable is shown as above </u>: ** 

  + Country is factor
  + continent is also factor
  + year is int
  + lifeExp is numberic
  + pop is int


---
## Explore individual variables  

Pick at least one categorical variable and at least one quantitative variable to explore: 

  * #### What are possible values (or range, whichever is appropriate) of each variable?
  
```{r, results='hold'}
summary(gapminder)
```
country and continent are categorical variables and year,lifeExp,pop,gdpPercap are quantitative variable; the possible values of each variable are shown as above.  

If we only want check the information of one value at each time, here are some functions we can use: 

```{r, results='hide'}
attach(gapminder)  # when using attach(), onjects in the database can be accesed by simply giving names. instead of using object$value

###################################  quantitative variable   #########################################
range(year)   # it is shown that the range of "year" is from 1952 to 2007

####  we can also use min() and max() to check the range
min(year)   # 1952
max(year)   # 2007

### what about summary？ 
summary(year)

#################################### categorical variable #########################################
summary(country) # unlike quantitative variable, summary(categorical variable) returns the frequenct of each category

### besides summary(), table() is also a nice try
table(country)

```


  * ####  What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
  
```{r,results='hold'}
summary(lifeExp)
table(continent)
```
  
**Answer: the typical values of lifeExp are: mean = 59.47 Median = 60.71; The overall spread of lifeExp is max-min = 1318939990, IQR = 3rd-1st = 16796000; There are 624 information from Africa, 300 from Americas, 396 from Asia, 360 from Europe and 24 from Oceania**

---
## Explore various plot types  

```{r,results='hide'}
### There is a lot of data, but I only want to foucs on the information from Asia

info_Asia <- gapminder %>% 
  filter(continent == 'Asia')
summary(info_Asia$country)
info_Asia <-rename(info_Asia, cou = country)
ggplot(info_Asia, aes(cou, pop)) +
  geom_boxplot(aes(fill = cou)) +
  ggtitle("Distrubtion of pop in Asia")
```

***There are too many countries in my graph, but I only have intrest in country with huge population***

```{r, results='hide',warning=FALSE}

info_Asia <- gapminder %>% 
  filter(continent == 'Asia' & pop > 1000000000)
summary(info_Asia$country)
info_Asia <-rename(info_Asia, cou = country)
ggplot(info_Asia, aes(cou, pop)) +
  geom_boxplot(aes(fill = cou)) +
  ggtitle("Distrubtion of pop in Asia")
```


**According to the garph above, we can find that China and India are the countries with large population in Asia**


***Now I want to get more information of China and India, and find some differences***

```{r,warning=FALSE}
China_India <- gapminder %>% 
  filter(country %in% c("China","India"))
dim(China_India)  ## check the size of China_India
knitr::kable(China_India) ## use table presentation
```

**I would to use tabke to have a first look and it seems that there are somthing intresting among population, year and gdp. Let's check it!**

```{r,warning=FALSE}
#### take China as an example
China <- gapminder %>% 
  filter(country == 'China') %>% 
  arrange(year,pop)
#knitr::kable(China)

ggplot(China, aes(pop, gdpPercap)) +
  geom_point(alpha = 0.5,colour = "blue") +
  ggtitle("Relationship between population and gdp")

ggplot(China, aes(year, gdpPercap)) +
  geom_point(alpha = 0.5,colour = "blue") +
  ggtitle("Relationship between year and gdp")
```









**As I guessed before, gdp grows up with year and population**

***if we want to see the all relationships between variables, use cool pairs()***

```{r}
pairs(gapminder)
```







***let's try put rainbow on garph***

```{r}
myData <- gapminder %>% dplyr::select(lifeExp)
binwidth = 0.5
library(ggplot2)   # CRAN version 2.2.1 used
n_bins <- length(ggplot2:::bin_breaks_width(range(myData), width = binwidth)$breaks) - 1L
ggplot() + 
  geom_histogram(aes(x = myData), binwidth = binwidth, fill = rainbow(n_bins)) +
  ggtitle("try rainbow()")

```


***one graph first and then put on another***

```{r}
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram(aes(y =..density..),fill = "blue",alpha = 0.1)+
  geom_density(color = "red")
```


```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot(aes(year, pop)) +
  geom_jitter( colour = "green")+
  geom_violin(colour ="red")

```

---
## Use filter(), select() and %>%
I used these functions in my hw02, expecially in **Explore various plot types**

 
 

 
---
## But I want to do more!

```{r}
afterFilter <- filter(gapminder, country == c("Rwanda", "Afghanistan"))
unique(afterFilter$country)
```
**Answer: yes it is a correct way**
